Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Put control flow back a little, fixed up properties

\newcommand{\powerset}{\raisebox{.15\baselineskip}{\Large\ensuremath{\wp}}}
\newcommand*{\rsp}{\textsc{sp}}

@@

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}\BCP{I
  gave it a name different from WBCF, but I could be talked out of that.  In
any case, it now needs to be discussed / defined in several places below.}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.).

% This approach is novel in that the
% confidentiality\bcp{only?} property applies to many nested sub-traces within a single execution,
% as opposed to traces of entire executions.
% {\em Integrity} says that each caller's local data is protected from
% modification by callees,
% motivated by the reasoning principle that local data persists across a
% call.

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em observational} variants.  The simpler {stepwise}
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The more extensional observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.
\bcp{Please check: is that accurate?}

@@

\bcp{I trimmed away a ton of text here that I thought was getting too much
  into the weeds for the introduction.  However, in the process I lost the
  explanation of how our properties are a bit stronger than LSE.  One way to
put it back would be simply to add to the contributions a claim that we are
a bit stronger than them, with a pointer to where we discuss this in detail.}

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security to formalize a {\em stepwise stack-safety}
property and argue that it captures the intuitive concept of stack
safety~(\cref{sec:prelim,sec:lse}). We show that the definition can be
extended to argument passing on the stack and to a simple coroutine system.
\item We use property-based randomized testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
\apt{I think this is worth counting as a separate contribution.}
\end{itemize}

@@

Our stack-safety properties depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code%
%
\footnote{\bcp{Some possibly useful text: A major challenge in
    characterizing stack safety formally is that, in a low-level machine
    without built-in stacks, we are at the mercy\bcp{?} of
    architectural\bcp{?}  choices that do not relate to the abstract notion
    of stack safety. Some implementations may have only a stack pointer
    while others add a frame pointer, in some systems return addresses and
    data live in separate stacks, and so on. We abstract over these choices
    as much as possible.  }}%
%
---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, another state fulfils its return target under a standard calling convention if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

\apt{Suddenly we are talking about alternative enforcement mechanisms for the first time.
  This surely needs \emph{some} foreshadowing in the abstract and intro?}
Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

@@

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation\bcp{Can we align the C-like snippets with
    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
      at places in both the C and asssembly, making them harder to read
  and eating more space.}}  \ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

We show our examples both in C syntax and in corresponding assembly code for a
simple two-address machine. The address of each instruction is shown to its left.
The assembly code reflects a simple compilation model in which
local variables are stored in the stack frame.
For the moment, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \%sp} points to the first unused word above the top of the stack.
The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
implicitly adjusting the stack pointer.
At the right of each example, we show the layout of stack memory just before
the specified instruction is executed; each box represents a memory word, and
addresses increase to the right.

@@

Meanwhile, confidentiality intuitively states that a callee is insensitive to 
the context of its call (except for its arguments and any global
variables).
We state this as a kind of {\em noninterference} property~\citep{??}.
Noninterference describes knowledge: it states that for system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

@@

This form of confidentiality we will term {\em stepwise confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. To step in lockstep
means that if both states step---that is, neither the primary nor the variant failstops---any
{\em system component} (such as memory cell, register, program counter) that changes
in one step has a matching value at the end of the other step. When we later
extend the model with a notion of observable output, values that are emitted will need
to match as well.

@@

\paragraph*{Control-Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}{.2\textwidth}
{\footnotesize
\begin{verbatim}
int *stash = 0;

@@

void f() {
  int y;
  if (!stash) {
    stash = *(&y-1);
  } else {
    *(&y-1) = stash;
    stash = 0;
  }
  return;
}
\end{verbatim}
}
\end{subfigure}
  \begin{subfigure}{.4\textwidth}
{\footnotesize
\begin{verbatim}
   stash: .word 0
1  main:  add $1,%sp     ; alloc frame
2         mov $1,-1(%sp) ; init x
3         call f
4         neg -1(%sp)    ; x = -x
5         call f
6         mov -1(%sp),%ra ; set ret val
7         sub $1,%sp     ; dealloc frame
6         ret

10 f:     add $1,%sp     ; alloc frame
11        cmp stash, $0  ; stash = 0? 
12        breq #15       ; branch if so
13        mov -2(%sp),stash ; set stash
14        jmp #17
15        mov stash,-2(%sp) ; get stash
16        mov $0,stash
17        sub $1,%sp     ; dealloc frame
18        ret
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.25\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \hspace*{-1.5em}
      \memory[1.2em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-2.2em}{0.75em}{\#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A control-flow attack}
  \label{fig:controlflow}
\end{figure}

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets of calls and returns nested within on another.
Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}.
Here we show return addresses explicitly as line numbers, in the form $\#n$.
Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
up being called a third time, at which point it finally returns to instruction 5.

@@

@@

@@

\judgmenttwo[ Call]
            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

\judgmenttwo[ ReturnFound]
            {\(\mach \stepsto \mach' \in \target\)}
            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}

\judgment[ Default]
         {}
         {\(\_,\context \stepsto_C \context\)}

\definition Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
Let \(\components\) be the set of components \(\component\) where
\(\domMap ~ \component = \sealed{\_}\) or \(\passed{\depth,\depth'}\) where \(|\rts| < \depth\)
or \(|\rts| > \depth'\). A system enjoys {\em stepwise stack integrity with passing} if for all
such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\).

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \rts)\) hold for
\(|\rts| \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\). Let
\(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \unsealed\) or \(\passed{\depth,\depth'}\) where
\(\depth \leq |\rts| \leq \depth'\) .
Then take any \(\nach\) such that \(\mach \approx_\components \nach\), and
\(\nach \hookrightarrow \NPCT | \neg f_\depth\). A system enjoys
{\em stepwise stack confidentiality with passing} if for all such \(\MPCT\) and \(\NPCT\),
\(\MPCT \doteq \NPCT\).

>>>>

\newcommand*{\rsp}{\textsc{sp}}

@@

In the present work, we also aim to protect both stack data (like
LSE) and
control flow (like WBCF).  But our formulations of these properties are
rather different from those of Skorstengaard et al.  Instead of
defining stack safety implicitly,
as ``what a certain idealized machine does,'' we define it
\emph{explicitly} using the technical framework of language-based
security\ifaftersubmission~\citep{??}\bcp{Maybe the Sabelfeld and Myers
  survey?}\fi.
Our core properties are \emph{stepwise stack integrity},
\emph{stepwise stack confidentiality}, and \emph{return to caller}.
%
A key technical novelty in the definitions of confidentiality and integrity,
compared to standard formulations from the security literature, is that they
are ``nested'': {\em each} caller is guaranteed protection from its
immediate callees (which, in turn, need protection from their immediate
callees, etc.). They are also stronger than LSE; even in the absence of the
control flow guarantees of {\em return to caller}, they guarantee callers'
data safety in the face of control-flow attacks (see \cref{fig:control} and discussion).
This reflects a cleaner division between data protection and control-flow protection
than offered by LSE and WBCF.

For the confidentiality and integrity properties, we consider both {\em
  stepwise} and {\em observational} variants.  The simpler {stepwise}
variants formalize the intuition that a caller's local data is {\em never}
read or modified during a call. The more extensional observational variants
allow a callee to read from and write to their caller's stack frame, as
long as nothing that they read affects their observable behavior and nothing
they write affects the observable behavior of the caller after they return.

@@

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security to formalize a {\em stepwise stack-safety}
property and argue that it captures the intuitive concept of strict stack
safety~(\cref{sec:prelim,sec:lse}). We show that the definition can be
extended to argument passing on the stack and to a simple coroutine system.
\item We use property-based randomized testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
\end{itemize}

@@

Our stack-safety properties depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code%
%
\footnote{\bcp{Some possibly useful text: A major challenge in
    characterizing stack safety formally is that, in a low-level machine
    without built-in stacks, we are at the mercy\bcp{?} of
    architectural\bcp{?}  choices that do not relate to the abstract notion
    of stack safety. Some implementations may have only a stack pointer
    while others add a frame pointer, in some systems return addresses and
    data live in separate stacks, and so on. We abstract over these choices
    as much as possible.  }}%
%
---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, in a standard calling convention, another state fulfils its return target if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

@@

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation\bcp{Can we align the C-like snippets with
    the corresponding assembly?}  \ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

@@

We show our examples both in C syntax and in corresponding assembly code for a
simple two-address machine. The address of each instruction is shown to its left.
The assembly code reflects a simple compilation model in which
local variables are stored in the stack frame.
For the moment, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack.
The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
implicitly adjusting the stack pointer.
At the right of each example, we show the layout of stack memory just before
the specified instruction is executed; each box represents a memory word, and
addresses increase to the right.

@@

Meanwhile, confidentiality intuitively states that a callee is unaware\bcp{??}
of the context of its call (except for its arguments and any global
variables).
We state this as a kind of {\em noninterference} property~\citep{??}.
Noninterference describes knowledge: it states that for system components
containing secret values, we can substitute any values whatsoever without changing the results
of our computation. If there exists some value for which we see different results, then
in fact we can learn something about the initial state by observing those results, violating
confidentiality.

@@

This form of confidentiality we will term {\em stepwise confidentiality}, defined informally
as: for any state at the entry to a function, for any variant over the values in the stack,
the primary state and its variant will step in lockstep until they return. To step in lockstep
means that if both states step -- that is, neither the primary nor the variant failstops --
any {\em system component} (such as memory cell, register, program counter) that changes
in one step has a matching value at the end of the other step. And when we
extend the model with a notion of external output, values that are outputted will need
to match as well.

@@

\paragraph*{Control-Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}{.3\textwidth}
{\small
\begin{verbatim}
int *stash = 0;

@@

void f() {
  int y;
  if (!stash) {
    stash = *(&y-1);
  } else {
    *(&y-1) = stash;
    stash = 0;
  }
  return;
}
\end{verbatim}
}
\end{subfigure}
  \begin{subfigure}{.65\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $1,%sp        ; allocate frame
2         mov $1,-1(%sp)    ; initialize x
3         call f
4         neg -1(%sp)       ; x = -x
5         call f
6         mov -1(%sp),%ra   ; set return value
7         sub $1,%sp        ; deallocate frame
6         ret

10 f:     add $1,%sp        ; allocate frame
11        cmp stash, $0     ; compare
12        breq #15          ; branch if equal
13        mov -2(%sp),stash ; update stash
14        jmp #17
15        mov stash,-2(%sp) ; retrieve stash
16        mov $0,stash
17        sub $1,%sp        ; deallocate frame
18        ret
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.4\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \memory[3em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A control-flow attack \bcp{wonder if we can cram the bottom part
      onto the same row as the rest.  Might need to make things smaller,
      spill into the margin, etc., but we really need the space.  (Another
      option is to make it a separate figure and flow text around it, but
      this will save less space.)}}
  \label{fig:controlflow}
\end{figure}

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets calls and returns nested within on another.
Rather, they aim to capture what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}.
Here we show return addresses explicitly as line numbers, in the form $\#n$.
Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
up being called a third time, at which point it finally returns to instruction 5.

@@

\vspace*{-1ex}
\judgmentthree[ReturnRec]
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
              {\(\components = \{\component | \domMap ~ \component = \sealed{|\rts|}\}\)}
              {\(\domMap'' = \domMap'[\components \mapsto \unsealed]\)}
              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}

@@

\subsection{Control Flow and Return-to-Caller}
\label{sec:RTC}

Together, integrity and confidentiality protect a caller's data until its callee returns
to it. But it is possible that an intermediate caller is never returned to, but returned
{\em past}, to a deeper caller. Fundamentally, this is a control-flow problem, not a
data-protection problem. From that caller's perspective, it is no different from the
callee diverging. But a lower level caller might have been relying on the intermediate
caller to finish its computation, so it is worth protecting control flow even from
attacks that do not violate data protections. We formalize this as {\em return-to-caller}.

\definition {\bf Return-To-Caller}

A system enjoys {\em return-to-caller} when, for every pair of reachable states
\(\mpcstatename = (\_,\_,(\_,\rts))\) and
\(\mpcstatename' = (\_,\_,(\_,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), \(|\rts| - 1 \leq |\rts'| \leq |\rts| + 1\).

@@

\judgmenttwo[Call]
            {\(\codemap ~ (\mach ~ \PCname) = \callmap(P)\)}
            {\(\domMap' = \mathit{push} ~ \domMap ~ \sealcon ~ P ~ \mach ~ |\rts|\)}
            {\(\mach,(\domMap,\rts) \stepsto_C \domMap',(\retcon ~ \mach)::\rts\)}

\vspace*{-1ex}
\judgmenttwo[ ReturnFound]
            {\(\mach \stepsto \mach' \in \target\)}
            {\(\domMap' = \mathit{pop} ~ \domMap ~ |\rts|\)}
            {\(\mach,(\domMap,\target::\rts) \stepsto_C \domMap',\rts\)}

\vspace*{-1ex}
\judgmentthree[ReturnRec]
              {\(\mach,(\domMap,\rts) \stepsto_C \domMap',\rts'\)}
              {\(\domMap'' = \mathit{pop} ~ \domMap' ~ |\rts|\)}
              {\(\mach,(\domMap,\_::\rts) \stepsto_C \domMap'',\rts'\)}

\definition Let the protected set of a context, \(\mathit{protected}(\domMap,\rts)\),
be the set of components \(\component\) such that either \(\domMap ~ \component = \sealed{\_}\)
or \(\domMap ~ \component = \passed{\depth,\depth'}\) and \(|\rts| < \depth\) or
\(|\rts| > \depth'\).

\definition {\bf Stepwise Stack Integrity (Passing)}

Let \(\mpcstatename\) and \(\mpcstatename'\) be reachable states such that
\(\mpcstatename \stepstocon \mpcstatename'\), and \(\pi_\context(\mpcstatename) = \domMap,\rts\).
A system enjoys {\em stepwise stack integrity with passing} if for all
such states, \(\pi_\mach(\mpcstatename) ~ \component = \pi_\mach(\mpcstatename') ~ \component\)
for each component in \(\mathit{protected}(\pi_\context(\mpcstatename))\)

\definition {\bf Stepwise Stack Confidentiality (Passing)}

For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,(\domMap,\rts)\),
let \(\components\) be the set of components \(\component\) such that
\(\domMap ~ \component = \outside\) or \(\domMap ~ \component = \passed{\depth,\depth'}\)
and \(\depth \leq |\rts| \leq \depth'\). A system enjoys
\emph{stepwise stack confidentiality} if,
for all \(\nach\) such that \(\mach \approx_\components \nach\) and
\(\nach \hookrightarrow \NPCT | \neg f_{|\rts|}\),
\(\MPCT \doteq \NPCT\).
